---
output:
  github_document:
      df_print: 'kable'
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#> ",
    fig.path = "man/figures/README-"
    # out.width = "100%"
)
```

# moexer

<!-- badges: start -->

<!-- badges: end -->

Moscow Exchange (MOEX) provides a REST interface to its Informational and Statistical Server (ISS), see <https://fs.moex.com/files/8888>.

`moexer` is a thin wrapper around the REST interface. It allows to quickly fetch e.g. price candles for a particular security, obtain its profile information and so on. The data is returned as `tibble`s, making it easy to subsequently process and analyse it.

## Installation

You can install the released version of moexer from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("moexer")
```

or the latest version from github:

``` r
devtools::install_github("x1o/moexer")
```

## High-Level Interface to ISS

Suppose you want to download monthly candles from Jan 1, 2020 until the present day for the Sberbank common shares.

```{r message=FALSE}
library(moexer)
library(dplyr)
library(purrr)
```

Each security on MOEX has an ID, e.g. a 4-letter ticker symbol for a share ("`LKOH`") or an ISIN for a bond ("`RU000A0JXPU3`").

Find the corresponding security ID:

```{r}
search_security(query = 'Sberbank') |> 
    slice_head(n = 10) |> 
    select(secid, name, is_traded, type, primary_boardid)
```

We can verify that `SBER` is indeed the symbol we were looking for and check the profile information:

```{r}
sber_info <- get_security_info(secid = 'SBER')
sber_info$description |> 
    select(name, title, value)
sber_info$boards |> 
    slice_head(n = 10) |> 
    select(secid, boardid, title, is_traded, history_from, history_till, currencyid)
```

Fetch the `SBER` candles:

```{r}
get_candles(secid = 'SBER', from = '2020-01-01', till = '2022-01-01', interval = 'monthly') |> 
    head()
```

`get_candles()` also supports specifying date-times for the `from`, `till` arguments,
e.g. `'2020-01-01 09:00:00'`.  This is most useful for obtaining intraday candles,
e.g with `interval = 'hourly'` --- see `options('moexer.candle.intervals')`.

If `till = NULL`, all candles up to today are fetched.

`get_candles()` is vectorised over `secid`, so it is possible to, say, fetch candles for both the common and the preferred shares. The returned object has class `MoexCandles` for which there's an appropriate `plot()` method:

```{r}
get_candles(
    secid = c('SBER', 'SBERP'), 
    from = '2020-01-01', 
    till = '2022-01-01', 
    interval = 'monthly'
) |> 
    plot()
```

## Low-level Interface to ISS

### Request

ISS is accessed via HTTP and/or HTTPS.

The interface is "RESTful", meaning the endpoint parameters can be passed as a query string, but they form the path of the URL. For instance, to receive `SBER` candles, one would need to form a URL as below:

```         
<base_url>/securities/SBER/candles?from=2020-01-10&till=2020-01-17
```

For a list of all public endpoints, see <http://iss.moex.com/iss/reference/>.

### Response

ISS is capable of return data in several formats; the present package uses JSON internally.

Every response consists of named sections ("blocks").

Every block contains the following sub-sections:

-   `metadata` with types
-   `column` with column names
-   `data` with the actual payload

All response blocks are parsed as tibbles using the information above.

### Query String Parameters

The parameters can be applied to some sections (1), all section (2) or modify "system" ISS parameters (3).

1.  `<block>.<parameter>`=`<value>` (applicable for block `<block>`)
    -   `<block>.columns`=`<id_1>,<id_2>,...`: only select these column in the block `<block>`
2.  `<parameter>`=`<value>` (all response blocks)
3.  `iss.<parameter>`=`<value>` (system parameter)
    -   `iss.only`=`<block_1>,<block_2>,...`: only return these blocks

### Examples

The main function for working with low-level requests is `query_iss()`.

#### Simple Queries

Fetch possible values for certain market objects. This returns a list of sections such as `engines`, `markets`, etc, each being a dataframe.

```{r}
iss_index <- query_iss('index')
names(iss_index)
```

```{r}
iss_index$engines
```

Query string parameters are specified as `params` argument; `debug_output = TRUE` results in the request URL being printed:

```{r}
df <- query_iss(
    'index',
    params = list(
        iss.only = 'engines,markets', 
        markets.columns = 'id,market_name,market_title'
    ),
    debug_output = TRUE
)
df$engines |> head()
df$markets |> head()
```

```{r}
show_df <- function(df) {
    print(nrow(df))
    bind_rows(head(df), tail(df))
}
```

#### Queries Involving Cursor

Some queries return large responses, which are split into pages that must accessed using `start` query string parameter as indicated in the auxiliary section `history.cursor`, also returned by ISS:

```{r}
df <- query_iss(
    'history/engines/stock/markets/shares/securities/MOEX',
    params = list(
        from = '2021-09-01',
        till = '2021-12-31',
        start = 10
    )
)
show_df(df$history[,1:10])
df$history.cursor
```

An adverb `following_cursor()` decorates `query_iss()` to follow the cursor until the requested information is received completely. The `.cursor` dataframe is dropped.

```{r}
following_cursor(query_iss)(
    'history/engines/stock/markets/shares/securities/MOEX',
    params = list(
        from = '2021-09-01',
        till = '2021-12-31',
        start = 10
    )
) |> 
    pluck('history') |> 
    select(1:10) |> 
    show_df()
```

#### Queries With Automatical `start` Increment

Some responses are large yet are not accompanied by a `.cursor` block.

```{r}
query_iss(
    'engines/stock/markets/shares/boards/TQBR/securities/SBER/candles',
    params = list(
        from = URLencode('2020-01-10 10:00:00'),
        till = URLencode('2020-01-10 23:59:59'),
        interval = 1,
        start = 10
    ),
    debug_output = TRUE
) |> 
    pluck('candles') |> 
    show_df()
```

These queries may accessed by `query_iss()` wrapped into `fetching_fully` decorator, which automatically increases the value of `start` parameter until no more data is received:

```{r}
fetching_fully(query_iss)(
    'engines/stock/markets/shares/boards/TQBR/securities/SBER/candles',
    params = list(
       from = URLencode('2020-01-10 10:00:00'),
       till = URLencode('2020-01-10 23:59:59'),
       interval = 1,
       start = 10
    )
) |> 
    pluck('candles') |> 
    show_df()
```

Some endpoints do not support `start` parameter -- `fetching_fully()` knows how to handle that:

```{r}
fetching_fully(query_iss)('turnovers', params = list(iss.only = 'turnovers'))$turnovers
```
