---
output:
  github_document:
      df_print: 'kable'
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#> ",
  fig.path = "man/figures/README-"
  # out.width = "100%"
)
```

# moexer

<!-- badges: start -->
<!-- badges: end -->

Moscow Exchange (MOEX) provides a REST interface
to its Informational and Statistical Server (ISS), see https://fs.moex.com/files/8888.

`moexer` is a thin wrapper around the REST interface.  It allows to quickly 
fetch e.g. price candles for a particular security, obtain its profile 
information and so on.  The data is returned as `tibble`s, making it easy to 
subsequently process and analyse it.

## Installation

You can install the released version of moexer from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("moexer")
```

or the latest version from github:

``` r
devtools::install_github("x1o/moexer")
```

## Example

Suppose you want to download monthly candles from Jan 1, 2020 until the present 
day for the Sberbank common shares.

```{r message=FALSE}
library(moexer)
library(dplyr)
```

Each security on MOEX has an ID, e.g. a 4-letter ticker symbol for a share
("LKOH") or an ISIN for a bond ("RU000A0JXPU3").

Find the corresponding security ID:

```{r}
search_security(query = 'Sberbank') |> 
    slice_head(n = 10) |> 
    select(secid, name, is_traded, type, primary_boardid)
```

We can verify that `SBER` is indeed the symbol we were looking for and check the profile information:

```{r}
sber_info <- get_security_info(secid = 'SBER')
sber_info$description |> 
    select(name, title, value)
sber_info$boards |> 
    slice_head(n = 10) |> 
    select(secid, boardid, title, is_traded, history_from, history_till, currencyid)
```

Fetch the `SBER` candles:

```{r}
get_candles(secid = 'SBER', from = '2020-01-01', interval = 'monthly')
```

`get_candles()` is vectorised over `secid`, so it is possible to, say, fetch 
candles for both the common and the preferred shares:

```{r}
get_candles(secid = c('SBER', 'SBERP'), from = '2020-01-01', interval = 'monthly') |> 
    plot_candles()
```

